Automatic Amortized Analysis for Imperative Code

### 2014-02-27 main.tex
try difference rules

We start with a programming language with loops counting
downwards to 0.  Statements and expressions are defined
below.  Assignments of complex expressions can be simulated
by sequencing assignments of simple expressions.
%
\begin{mathpar}
S := \code{assert}(\Gamma)
\mid x \leftarrow E
\mid \code{while}(0 < x)~S
\mid S;S

@@

E := x + y \mid x - y % \mid y
\end{mathpar}
%
We define a deductive system to derive judgements of the
form $\Gamma; Q \vdash S \dashv Q'; \Delta$ with the following
intended semantics
$$
\forall H_1\, H_2\, S\, n,
    ( H_1 \models \Gamma \land
      (H_1, S) \Downarrow_n H_2
    ) \Rightarrow
    ( H_2 \models \Delta \land
      \Phi_{H_1}(Q) \ge n + \Phi_{H_2}(Q')
    ).
$$

@@

\Rule{Assert}
{}
{ \Gamma; Q \vdash \code{assert}(\Delta) \dashv Q; \Gamma, \Delta }

\\ \Rule{AssignP}
{ \Gamma \models \Delta[x/x+y] }
{\Gamma; Q + y\cdot q_x \vdash x \leftarrow x + y \dashv Q; \Delta}

\and\Rule{AssignM}
{  \Gamma \models \Delta[x/x-y]
\\ \Gamma \models c \le y \le x
}
{ \Gamma; Q \vdash x \leftarrow x - y \dashv c \cdot q_x + V; \Delta }

\\ \Rule{While}
{ \Gamma, 0 < x ; Q \vdash S \dashv Q; \Gamma }
{ \Gamma; Q + M_w \vdash \code{while}(0 < x)~S \dashv Q; \Gamma }

\and\Rule{Seq}
{  \Gamma; Q \vdash S_1 \dashv Q'; \Delta
\\ \Delta; Q' \vdash S_2 \dashv Q''; \Gamma'
}
{ \Gamma; Q \vdash S_1;S_2 \dashv Q''; \Gamma' }

@@

\section{Decreasing differences}
The second iteration is a system more symmetric where the potentials
are linear functions in differences of variables.  This time, the while
condition can be a bit more complex.  In the rest, $k$
denotes a non-negative constant.
$$
S := \code{while}(k < x - y) \mid \dots
$$

>>>>

We start with a programming language with loops counting
downwards to 0.  Statements and expressions are defined
below.  Assignments of complex expressions can be simulated
by sequencing assignments of simple expressions.
%
\begin{mathpar}
S := \code{skip}
\mid \code{assert}(\Gamma)
\mid x \gets E
\mid \code{while}(0 < x)~S
\mid \code{if}(?)~S~\code{else}~S
\mid S;S

@@

E := x + y \mid x - y % \mid y
\end{mathpar}
%
We define a deductive system to derive judgements of the
form $\Gamma; Q \vdash S \dashv Q'; \Gamma'$ with the following
intended semantics
$$
\forall H\, H'\, S\, n,
    ( H \models \Gamma \land
      (H, S) \Downarrow_n H'
    ) \Rightarrow
    ( H' \models \Gamma' \land
      \Phi_{H}(Q) \ge n + \Phi_{H'}(Q')
    ).
$$

@@

\Rule{Skip}
{}
{ \Gamma; Q \vdash \code{skip} \dashv Q; \Gamma }

\and \Rule{Assert}
{}
{ \Gamma; Q \vdash \code{assert}(\Delta) \dashv Q; \Gamma, \Delta }

\\ \Rule{AssignP}
{ \Gamma \models \Gamma'[x/x+y] }
{\Gamma; Q + y\cdot q_x \vdash x \gets x + y \dashv Q; \Gamma'}

\and\Rule{AssignM}
{  \Gamma \models \Gamma'[x/x-y]
\\ \Gamma \models c \le y \le x
}
{ \Gamma; Q \vdash x \gets x - y \dashv c \cdot q_x + V; \Gamma' }

\\ \Rule{While}
{ \Gamma, 0 < x ; Q - M_w \vdash S \dashv Q; \Gamma }
{ \Gamma; Q \vdash \code{while}(0 < x)~S \dashv Q; \Gamma, x \le 0 }

\and\Rule{Seq}
{  \Gamma; Q \vdash S_1 \dashv Q'; \Gamma'
\\ \Gamma'; Q' \vdash S_2 \dashv Q''; \Gamma''
}
{ \Gamma; Q \vdash S_1;S_2 \dashv Q''; \Gamma'' }

@@

\section{Decreasing differences}
The second iteration is a system more symmetric where the potentials
are linear functions in differences of variables.  From now on $k$
denotes a non-negative constant.
$$
S := \code{while}(k < x - y)~S \mid \dots
$$

\Rule{AssignOpK}
{  \Gamma \models \Gamma'[x/x \pm y], y = k
\\ \Gamma \models x \pm y \in [x, u]
\\ \textstyle q'_y = q_y + \sum_u q^1_{xu} - \sum_v q_{xv}
\\ q_{xu} = q^1_{xu} + q^2_{xu}
}
{ \textstyle
  \Gamma; Q \vdash x \gets x \pm y \dashv \sum_u k \cdot q^2_{xu} + Q'; \Gamma'
}

\and \Rule{AssignOp}
{  \Gamma \models \Gamma'[x/x \pm y]
\\ \Gamma \models x \pm y \in [x, u]
\\ \textstyle q'_y = q_y + \sum_u q_{xu} - \sum_v q_{xv}
}
{ \textstyle
  \Gamma; Q \vdash x \gets x \pm y \dashv Q'; \Gamma'
}

\\ \mbox{(The other rules are unchanged)}
